---
title: "Untitled"
output: html_document
---

```{r}
library(tidyverse)
library(data.table)
library(readr)
library(permuco)
library(writexl)
library(readxl)
```
##HLA data
hla_table.RData == FID merged (by Shishir) with 2-digit imputed HLA from Greg

##Phenotypes

##load data object into R workspace
```{r}
load("/mnt/share6/SHARED_DATASETS/hla_working_group/results/hla_table.RData")

hla_table <- hla_table %>% 
  mutate(FID = as.character(FID))

unique_id_num <-  hla_table %>% 
  distinct(FID) %>% 
  nrow()

nrow(hla_table) == unique_id_num
```


##load phenotypes
```{r}
pheno = fread("/mnt/share6/For_Takeo/phenotypdata/pancreas_subclinical_phenotypes.txt")
pheno <- pheno %>% 
  mutate(FID = str_replace_all(GeneticID,"-","0"))

pheno <- pheno %>% 
  select(FID,pancreatitis,Diag)

pheno <- pheno %>% 
  mutate(pancreatitis = ifelse(pancreatitis == 1,0,
                               ifelse(pancreatitis == 2,1,NA)))

```
add [colClasses = "character"] to preserve, if genetic-ID is dropping leading zeroes -- read.table("filename.txt", header = TRUE, colClasses = "character")

```{r}
dim(pheno)
head(pheno)
```
if need to check header and dimensions




###convert pheno to numeric AS NEEDED
```{r}
for (i in 2:ncol(pheno)){
  pheno[,i] <- as.numeric(pheno[,i])
}
```



##load PCs, if including as covariates and not included in text file alongside phenotypes
```{r}
pc <- fread("/mnt/share6/FOR_Takeo/ICHIP/PC_ichip1to8washubbc_cauc.txt")
```

## add pc and phenotype information to hla_table

```{r}

hla_v2 <- hla_table %>% 
  left_join(pheno, by = "FID")

hla_v3 <- hla_v2 %>% 
  left_join(pc,by = "FID")


hla_v3 <- hla_v3 %>% 
  filter(!is.na(pancreatitis))
```


## remove relatedness
```{r}
relate = fread("/mnt/isilon_data/For_Takeo2/From_nick/Ichip/related_ichip")


hla_v3 <- hla_v3 %>% 
  filter(!FID %in% relate$FID) 


hla_v3 <- hla_v3 %>% 
  filter(!is.na(PC1))
  
```
## remove umbigous phenotype subjects
```{r}

hla_v3 <- hla_v3 %>% 
  filter(Diag %in% c("CD","UC","IBDU"))

```




```{r}
AJ = fread("/mnt/share6/FOR_Takeo/hla/hla_AJ_list")
non_AJ = fread("/mnt/share6/FOR_Takeo/hla/hla_nonAJ_list")


hla_aj = hla_v3 %>% 
  filter(FID %in% AJ$FID)

hla_nonaj = hla_v3 %>% 
  filter(FID %in% non_AJ$FID)


dim(hla_v3)

```



##Logistic regression with 4 PCs 
From this point, calculate logistic pvlue and OR of each HLA genotype for each phenotype.
Add 4 PCs as covariates.
Note that if logistic model give you NA p value (in my case, case and contorols are skewed, so
it can happen that no cases in alt allele pattern), just using coef(test) will not return phenotype pvalue.
You may accidentaly get pvalue for PC1 instead of phenotype. 
To avoid this phenomenon, I used full_join method 
(https://stackoverflow.com/questions/48818658/r-extract-glm-coefficients-including-na-rows). 



### prepare out put file

```{r}
kotae_nonaj <- data.frame(x = c(1:104))
```

### make vector of names of your phenotype
```{r}
pheno_name <- "pancreatitis"


```


```{r}

for (j in 1:length(pheno_name)) {
  pheno_1 <- pheno_name[j]
  pheno1_log_regr <- data.frame()
  
  for(i in 2:105) {
    logreg <- paste(pheno_1, "~", paste(names(hla_nonaj)[i], "+ PC1 + PC2 + PC3 + PC4"))
    fit <- glm(as.formula(logreg), data=hla_nonaj, family="binomial")
    test <- summary(fit)
    # this part is to avoid getting pvalue of PC1 by mistake
    tempy <- data.frame(coef(test)) %>%
      rownames_to_column("variable") %>% 
      full_join(data.frame(variable = attr(test$terms, "term.labels"))) %>% 
      arrange(variable) %>% 
      filter(grepl("HLA",variable))
    
    pheno1_log_regr[i-1,1] <-exp(tempy[1,2])
    pheno1_log_regr[i-1,2] <-tempy[1,5]
    pheno1_log_regr[i-1,3] <- names(hla_nonaj)[i]
   }
  
  p_name <- paste0(pheno_1,"_P")
  or_name <- paste0(pheno_1,"_OR")
  pheno1_log_regr <- pheno1_log_regr %>% 
    dplyr::rename(!!or_name := V1) %>% 
    dplyr::rename(!!p_name := V2) %>% 
    dplyr::rename(!!paste0(pheno_1,"_allele") := V3)
  
kotae_nonaj <- cbind(kotae_nonaj,pheno1_log_regr)
}

```

## add maf info

```{r}
kotae_maf_nonaj = data.frame()
  
  for(i in 2:105) {
    allele_name = names(hla_nonaj)[i]
    # this part is to avoid getting pvalue of PC1 by mistake
    maf_df <- hla_nonaj %>% 
      rename(target := UQ(rlang::sym(allele_name))) %>% 
      select(pancreatitis,target)
    
    alt_num_cont = maf_df %>% filter(pancreatitis == 0 & target == 2) %>% nrow*2 +
      maf_df %>% filter(pancreatitis == 0 & target == 1) %>% nrow() 
    num_all_cont= maf_df %>% filter(pancreatitis == 0) %>% nrow()*2
    
    maf_cont = alt_num_cont/num_all_cont
    
    
    alt_num_case = maf_df %>% filter(pancreatitis == 1 & target == 2) %>% nrow*2 +
      maf_df %>% filter(pancreatitis == 1 & target == 1) %>% nrow() 
    num_all_case= maf_df %>% filter(pancreatitis == 1) %>% nrow()*2
    
    maf_case = alt_num_case/num_all_case
    
    
    tmp_kotae = data.frame(MAF_case = maf_case, MAF_cont = maf_cont, pancreatitis_allele = allele_name)
    
    
    
  
  kotae_maf_nonaj = rbind(kotae_maf_nonaj,tmp_kotae)
  }


```

```{r}
kotae_nonaj1 <- kotae_nonaj %>% 
  left_join(kotae_maf_nonaj,by = "pancreatitis_allele")
  
```



```{r}
kotae_aj <- data.frame(x = c(1:104))
```


```{r}
for (j in 1:length(pheno_name)) {
  pheno_1 <- pheno_name[j]
  pheno1_log_regr <- data.frame()
  
  for(i in 2:105) {
    logreg <- paste(pheno_1, "~", paste(names(hla_aj)[i], "+ PC1 + PC2 + PC3 + PC4"))
    fit <- glm(as.formula(logreg), data=hla_aj, family="binomial")
    test <- summary(fit)
    # this part is to avoid getting pvalue of PC1 by mistake
    tempy <- data.frame(coef(test)) %>%
      rownames_to_column("variable") %>% 
      full_join(data.frame(variable = attr(test$terms, "term.labels"))) %>% 
      arrange(variable) %>% 
      filter(grepl("HLA",variable))
    
    pheno1_log_regr[i-1,1] <-exp(tempy[1,2])
    pheno1_log_regr[i-1,2] <-tempy[1,5]
    pheno1_log_regr[i-1,3] <- names(hla_aj)[i]
   }
  
  p_name <- paste0(pheno_1,"_P")
  or_name <- paste0(pheno_1,"_OR")
  pheno1_log_regr <- pheno1_log_regr %>% 
    dplyr::rename(!!or_name := V1) %>% 
    dplyr::rename(!!p_name := V2) %>% 
    dplyr::rename(!!paste0(pheno_1,"_allele") := V3)
  
kotae_aj <- cbind(kotae_aj,pheno1_log_regr)
}

```


## add maf info

```{r}
kotae_maf_aj = data.frame()
  
  for(i in 2:105) {
    allele_name = names(hla_aj)[i]
    # this part is to avoid getting pvalue of PC1 by mistake
    maf_df <- hla_aj %>% 
      rename(target := UQ(rlang::sym(allele_name))) %>% 
      select(pancreatitis,target)
    
    alt_num_cont = maf_df %>% filter(pancreatitis == 0 & target == 2) %>% nrow*2 +
      maf_df %>% filter(pancreatitis == 0 & target == 1) %>% nrow() 
    num_all_cont= maf_df %>% filter(pancreatitis == 0) %>% nrow()*2
    
    maf_cont = alt_num_cont/num_all_cont
    
    
    alt_num_case = maf_df %>% filter(pancreatitis == 1 & target == 2) %>% nrow*2 +
      maf_df %>% filter(pancreatitis == 1 & target == 1) %>% nrow() 
    num_all_case= maf_df %>% filter(pancreatitis == 1) %>% nrow()*2
    
    maf_case = alt_num_case/num_all_case
    
    
    tmp_kotae = data.frame(MAF_case = maf_case, MAF_cont = maf_cont, pancreatitis_allele = allele_name)
    
    
    
  
  kotae_maf_aj = rbind(kotae_maf_aj,tmp_kotae)
  }

```

```{r}
kotae_aj1 <- kotae_aj %>% 
  left_join(kotae_maf_aj,by = "pancreatitis_allele")
```




```{r}
kotae_m = read_xlsx("/mnt/share6/FOR_Takeo/hla/thipurine_hla_pancreatitis.xlsx")
kotae_m <- kotae_m %>% 
  select(-x)

kotae_aj2 = kotae_aj1 %>% 
  select(-x) %>% 
  rename(OR_aj = pancreatitis_OR,
         P_aj = pancreatitis_P,
         MAF_case_aj = MAF_case,
         MAF_cont_aj= MAF_cont)

kotae_nonaj2 <- kotae_nonaj1 %>% 
  select(-x) %>% 
  rename(OR_nonaj = pancreatitis_OR,
         P_nonaj = pancreatitis_P,
         MAF_case_nonaj = MAF_case,
         MAF_cont_nonaj = MAF_cont)


kotae_m1 <- kotae_m %>% 
  left_join(kotae_aj2,by = "pancreatitis_allele") %>% 
  left_join(kotae_nonaj2,by = "pancreatitis_allele") %>% 
  rename(P_all = pancreatitis_P,
         OR_all = pancreatitis_OR,
         MAF_case_all = MAF_case,
         MAF_cont_all = MAF_cont) %>% 
  select(pancreatitis_allele,OR_all,P_all,MAF_case_all,MAF_cont_all,OR_aj,P_aj,MAF_case_aj,
         MAF_cont_aj,OR_nonaj,P_nonaj,MAF_case_nonaj,MAF_cont_nonaj)



```


## do conditional analysis with HLA_DRB1_7



```{r}
target_names = colnames(hla_v3)
target_names = target_names[grepl("HLA",target_names)]
target_names = target_names[!grepl("HLA_DRB1_7|HLA_DQA1_2",target_names)]

```



```{r}
kotae_conp <- data.frame(x = c(1:102))



for (j in 1:length(pheno_name)) {
  pheno_1 <- pheno_name[j]
  pheno1_log_regr <- data.frame()
  
  for(i in 1:length(target_names)) {
    logreg <- paste0(pheno_1, "~", target_names[i], "+ HLA_DRB1_7 + PC1 + PC2 + PC3 + PC4")
    fit <- glm(as.formula(logreg), data=hla_v3, family="binomial")
    test <- summary(fit)
    # this part is to avoid getting pvalue of PC1 by mistake
    tempy <- data.frame(coef(test)) %>%
      rownames_to_column("variable") %>% 
      full_join(data.frame(variable = attr(test$terms, "term.labels"))) %>% 
      arrange(variable) %>% 
      filter(grepl("HLA",variable))
    
    pheno1_log_regr[i,1] <-exp(tempy[1,2])
    pheno1_log_regr[i,2] <-tempy[1,5]
    pheno1_log_regr[i,3] <- target_names[i]
   }
  
  p_name <- paste0(pheno_1,"_P")
  or_name <- paste0(pheno_1,"_OR")
  pheno1_log_regr <- pheno1_log_regr %>% 
    dplyr::rename(!!or_name := V1) %>% 
    dplyr::rename(!!p_name := V2) %>% 
    dplyr::rename(!!paste0(pheno_1,"_allele") := V3)
  
kotae_conp <- cbind(kotae_conp,pheno1_log_regr)

print(i)
}




```

```{r}

kotae_conp1 <- kotae_conp %>% 
  select(-x,) %>% 
  rename(conditional_P = pancreatitis_P,
         conditional_OR = pancreatitis_OR)
kotae_m2 <- kotae_m1 %>% 
  left_join(kotae_conp1,by = "pancreatitis_allele") %>% 
  select(pancreatitis_allele,OR_all,P_all,conditional_OR,conditional_P,everything()) 

kotae_m2 %>% 
  write_xlsx("/mnt/share6/FOR_Takeo/hla/thipurine_hla_pancreatitis_aj_nonaj_rev.xlsx")
  
```


