---
title: "Untitled"
output: html_document
---


##HLA data
hla_table.RData == FID merged (by Shishir) with 2-digit imputed HLA from Greg

##Phenotypes

##load data object into R workspace
```{r}
load("/mnt/share6/SHARED_DATASETS/hla_working_group/results/hla_table.RData")

hla_table <- hla_table %>% 
  mutate(FID = as.character(FID))

unique_id_num <-  hla_table %>% 
  distinct(FID) %>% 
  nrow()

nrow(hla_table) == unique_id_num
```
n=12,904 subjects, with 104 alleles across HLA-2 digit
ls() to confirm has been loaded into environment

##load phenotypes
```{r}
pheno <- read_xlsx("/mnt/share6/FOR_Takeo/phenotypdata/Thiopurine toxicityforDalin_v2.xlsx") %>% 
  data.frame() %>% 
  mutate(FID = str_replace_all(LAB.ID,"-","0")) %>% 
  dplyr::select(FID,everything()) %>% 
  distinct(FID,.keep_all = TRUE) %>% 
  dplyr::select(-LAB.ID) %>% 
  dplyr::rename(Leukopenia = Leukopenia..Yes.1..No.0) %>% 
  dplyr::rename(GI.Intolerance = GI.Intolerance..N.V..abd.pain..diarrhea.) %>% 
  dplyr::rename(Allergy = Allergy..rash..high.fever..hypersensitivity.rxn.) %>% 
  dplyr::rename(Flu.like = Flu.like.sx..fever..malaise..nausea..HA.)



pheno[ifelse(pheno == "U",TRUE,FALSE)] <- NA
pheno[ifelse(pheno == "u",TRUE,FALSE)] <- NA


```
add [colClasses = "character"] to preserve, if genetic-ID is dropping leading zeroes -- read.table("filename.txt", header = TRUE, colClasses = "character")

```{r}
dim(pheno)
head(pheno)
```
if need to check header and dimensions

###convert pheno to numeric AS NEEDED
```{r}
for (i in 2:ncol(pheno)){
  pheno[,i] <- as.numeric(pheno[,i])
}
```



##load PCs, if including as covariates and not included in text file alongside phenotypes
```{r}
pc <- fread("/mnt/share6/FOR_Takeo/ICHIP/PC_ichip1to8washubbc_cauc.txt")
```

## add pc and phenotype information to hla_table

```{r}

hla_v2 <- hla_table %>% 
  left_join(pheno, by = "FID")

hla_v3 <- hla_v2 %>% 
  left_join(pc,by = "FID")

```



##Logistic regression with 4 PCs 
From this point, calculate logistic pvlue and OR of each HLA genotype for each phenotype.
Add 4 PCs as covariates.
Note that if logistic model give you NA p value (in my case, case and contorols are skewed, so
it can happen that no cases in alt allele pattern), just using coef(test) will not return phenotype pvalue.
You may accidentaly get pvalue for PC1 instead of phenotype. 
To avoid this phenomenon, I used full_join method 
(https://stackoverflow.com/questions/48818658/r-extract-glm-coefficients-including-na-rows). 



### prepare out put file

```{r}
kotae <- data.frame(x = c(1:104))
```

### make vector of names of your phenotype
```{r}
pheno_name <- colnames(pheno)[-1]
pheno_name

```


```{r}

for (j in 1:length(pheno_name)) {
  pheno_1 <- pheno_name[j]
  pheno1_log_regr <- data.frame()
  
  for(i in 2:105) {
    logreg <- paste(pheno_1, "~", paste(names(hla_v3)[i], "+ PC1 + PC2 + PC3 + PC4"))
    fit <- glm(as.formula(logreg), data=hla_v3, family="binomial")
    test <- summary(fit)
    # this part is to avoid getting pvalue of PC1 by mistake
    tempy <- data.frame(coef(test)) %>%
      rownames_to_column("variable") %>% 
      full_join(data.frame(variable = attr(test$terms, "term.labels"))) %>% 
      arrange(variable) %>% 
      filter(grepl("HLA",variable))
    
    pheno1_log_regr[i-1,1] <-exp(tempy[1,2])
    pheno1_log_regr[i-1,2] <-tempy[1,5]
    pheno1_log_regr[i-1,3] <- names(hla_v3)[i]
   }
  
  p_name <- paste0(pheno_1,"_P")
  or_name <- paste0(pheno_1,"_OR")
  pheno1_log_regr <- pheno1_log_regr %>% 
    dplyr::rename(!!or_name := V1) %>% 
    dplyr::rename(!!p_name := V2) %>% 
    dplyr::rename(!!paste0(pheno_1,"_allele") := V3)
  
kotae <- cbind(kotae,pheno1_log_regr)
}

```


### select necessary columns and make the table pretty
```{r}

kotae1 <- kotae %>% 
  dplyr::select(Leukopenia_allele,contains("_P"),contains("_OR")) %>% 
  dplyr::select(sort(names(.))) %>% 
  dplyr::select(Leukopenia_allele, everything()) %>% 
  dplyr::rename(Allele = Leukopenia_allele)

```


## Since my case-control rate is skewed, there were many warning messages for the logistic test,
## because of the genotype-phenotype table described above.
## To avoid this, I also did permutation tests for linear models.
## This will take more than one hour. 

### prepare out put file 
```{r}
kotae_perm <- data.frame(x = c(1:104))
```

```{r}
for (j in 1:length(pheno_name)) {
  pheno_1 <- pheno_name[j]
  pheno1_line_perm <- data.frame()
  
  for(i in 2:105) {
    linper <- paste(pheno_1, "~", paste(names(hla_v3)[i], "+ PC1 + PC2 + PC3 + PC4"))
    fit <- lmperm(as.formula(linper),data = hla_v3, np = 5000)
    
    tempy <- as.data.frame(fit$table)
    pheno1_line_perm[i-1,1] <-exp(tempy[2,1])
    pheno1_line_perm[i-1,2] <-tempy[2,7]
    pheno1_line_perm[i-1,3] <- names(hla_v3)[i]
  }
  
  p_name <- paste0(pheno_1,"_P_perm")
  or_name <- paste0(pheno_1,"_OR_perm")
  pheno1_line_perm <- pheno1_line_perm %>% 
    dplyr::rename(!!or_name := V1) %>% 
    dplyr::rename(!!p_name := V2) %>% 
    dplyr::rename(!!paste0(pheno_1,"_allele") := V3)
  
  kotae_perm <- cbind(kotae_perm,pheno1_line_perm)
}

```


## merge logistic result and permutation linear result.

```{r}

kotae_final <- kotae1 %>% 
  left_join(kotae_perm1,by = "Allele") %>% 
  dplyr::select(sort(names(.))) %>% 
  dplyr::select(-contains("OR_perm"))

```

# export result

```{r}
kotae_final %>% 
  write_xlsx("/mnt/share6/FOR_Takeo/hla/thipurine_hla_test.xlsx")
```

